Adopting agile software development in the large organizations: A systematic literature review

### 2014-06-02 journal/article.tex
T: Transformation process

This section presents the findings of the literature review. In the first
subsection a meta analysis of the analyzed papers is presented. In the
subsequent sections findings are presented organized according to the research
questions. The methods used for extracting the results were described in section
\ref{sec:method}. Note that the term \emph{study} is used when referring to a
primary study publication, and the term \emph{case} is used when referring to an
organization which may be described in several studies.

@@

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Business reasons for starting a change   &  Described in  \\
        \midrule
        Time to market must be reduced    &  P17, P18, P32, P38, P39, P47  \\
        Delivery is considered slow       &  P2, P8, P25, P36, P46, P49  \\
        Something must be done to remain competitive  &
                P9, P10, P11, P18, P19, P32, P38, P42, P49  \\
        Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
        Need to accommodate change during projects  &
                P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
        \bottomrule
    \end{tabular}
    \caption{Business related reasons for starting an agile transformation}
    \label{table:reasonstochange_business}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Management problems           &  Described in  \\
        \midrule
        Software development is managed badly   &  P8, P21, P30, P38, P44, P49  \\
        Command and control style management    &  P9, P25, P30, P40, P44  \\
        Scheduling and estimation is difficult  &  P16, P19, P21, P30, P33, P40, P41, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Management problems stated as reasons for organizational change}
    \label{table:reasonstochange_management}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Problems in the old process           &  Described in  \\
        \midrule
        Excess and overhead in process     &  P9, P24, P35, P36, P38, P40, P46, P49, P51  \\
        The process is being ignored       &  P6, P9, P46, P49  \\
        Late integration, testing, and feedback  &  P5, P25, P33, P40, P50  \\
        New practices can not be integrated in the old process  &  P10, P26  \\
        The old process does not scale up  &  P16, P30, P50, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Problems because of which the old process should be changed}
    \label{table:reasonstochange_process}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
        Reasons to change           &  Described in  \\
        \midrule
        Teams can not collaborate properly   &  P21, P38  \\
        Silos and internal boundaries in organization   &  P27, P40, P46  \\
        Lack of customer collaboration       &  P4, P26, P40  \\
        Demand for better quality            &  P2, P4, P25, P31, P30, P38, P41, P42  \\
        \bottomrule
    \end{tabular}
    \caption{Other reasons why organizations wanted to change the way of working}
    \label{table:reasonstochange_other}
\end{table}

@@

The second research question was aimed to discover how the process of
transformation can be described.
We provide an answer to Research Question 2 in several parts. First the initial
states of organizations are described. Secondly, we present a two-axial model
for categorizing transformation types. In the rest of this section we discuss
typical arrangements that relate to the change process, such as how training is
arranged, how change is lead, and what investments are done.

The challenge of giving a generic description on how transformations proceed is
that every organization is unique, and further, publications tended to provide
more focused viewpoints rather than describing the transformation process
itself. Therefore it was not sensible to attempt constructing a detailed model
on how large organizations adopt agile methods.
Table \ref{table:transformation_descriptiontypes} presents types of insights
each primary source provided in relation to Research Question 2.

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
        How transformation is described   &  Primary sources  \\
        \midrule
        Narrative description of transformation process &
                P3, P5, P6, P9, P11, P12, P16, P18,
                P24, P27, P30, P36, P40, P41, P46   \\
        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
                P10, P13, P15, P19, P23, P32, P34,
                P39, P47, P48, P50, P52    \\
        Viewpoints of challenges and success factors  &
                P17, P20, P21, P29, P37, P38, P49   \\
        Discussion on establishing large scale agile practices  &
                P7, P8, P31, P42, P43, P44, P45   \\
        Only brief discussion on transformation process  &
                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
        \bottomrule
    \end{tabular}
    \caption{The theme of each primary source, in relation to providing
             insights of the transformation process}
    \label{table:transformation_descriptiontypes}
\end{table}

@@

* Top down

* Bottom up

* Big bang

* Step wise

* Transformation is continuous, hard work, and takes time

Mentioning big bang versus step wise [P16].

@@

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Change leaders         &  Described in  \\
        \midrule
        High-up change leaders or sponsors  &
                P3, P5, P24, P31, P36, P38, P46, P49  \\
        Agile task group for coordinating change   &
                P4, P16, P18, P21, P23, P25, P30, P40, P41,
                P43, P45, P48, P49, P50  \\
        Change leaders on other organizational levels  &
                P2, P6, P9, P14, P17, P18, P27  \\
        Informal spreading of agile awareness  &
                P3, P6, P14, P40  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of entities leading change in the agile transformation}
    \label{table:transformation_leadership}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
        Activity           &  Described in  \\
        \midrule
        Training    &
                P3, P6, P8, P9, P11, P16, P18-P21, P23-P25, P27, P30,
                P32, P35, P37, P39-P41, P43-P46, P48-P51  \\  % 29
        Coaching    &
                P1-P3, P6, P8, P9, P11, P16, P18, P23, P27, P29,
                P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
        In-house coaching   &
                P1-P3, P6, P11, P16, P23, P29, P31, P32,
                P39, P41, P45, P47, P49, P50, P52  \\    % 17
        Seminars and workshops   &
                P1-P4, P7, P9, P10, P18, P11, P16, P23, P25,
                P26, P27, P32, P41, P44, P46, P47   \\   % 19
        Communities, special interest groups  &
                P3, P6, P7, P9, P10, P11, P12, P14, P32, P41,
                P42, P45, P47, P52  \\   % 14 
        \bottomrule
    \end{tabular}
    \caption{Summary of activities for building awareness and knowledgeability
             in the agile way of working}
    \label{table:transformation_training}
\end{table}

@@

A logical first step in an agile transformation is to try out the new methods in
a pilot project. A large amount of the primary studies described how piloting
was used in the beginning of the transformation. Primary sources discussing
the use of piloting are listed in \ref{table:transformation_piloting}.

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
        Sources discussing viewpoints on piloting   & 
                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
                P36, P38, P40, P41, P43, P45, P46, P47  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of activities aiming to advance the transformation}
    \label{table:transformation_piloting}
\end{table}

@@

Even if any activities aiming for transformation can be considered as
investments in themselves, the transformation process included some specific
tangible investments. These investments were focused on better enabling agile
software development, and highlight the concerns that organizations had when
implementing an agile model. Typical investments were bringing in external help,
new management tools, improvements in testing environments, and updating of the
physical working spaces. Table \ref{table:transformation_investments} summarizes
common types of investments.

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.35\textwidth}
                     >{\raggedright\arraybackslash}p{0.57\textwidth} }
        Investment for transformation      &  Described in  \\
        \midrule
        Using external consultants  &
                P5-P9, P16-P21, P23 P25, P27, P29-P31,
                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
        Agile project management tools  &
                P10, P11, P14, P16, P17, P26, P33,
                P41, P43, P44, P45, P48, P50 \\
        Continuous builds and testing  &
                P5, P11, P16, P17, P25, P27, P34, P39, P42, P50 \\
        Proprietary agile material or documentation   & 
                P1, P2, P8, P14, P16, P23, P26, P29, P32, P45  \\
        Refurbishing work areas and other spaces   &
                P6, P22, P35, P36, P38, P42, P46, P48 \\
        \bottomrule
    \end{tabular}
    \caption{Summary of investments aiming to make agile methods perform better}
    \label{table:transformation_investments}
\end{table}

>>>>

This section presents the findings of the literature review. In the first
subsection a meta analysis of the analyzed papers is presented. In the
subsequent sections findings are presented organized according to the research
questions.

The methods used for extracting the results were described in section
\ref{sec:method}. Note that the term \emph{study} is used when referring to a
primary study publication, and the term \emph{case} is used when referring to an
organization which may be described in several studies.

@@

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        \toprule
        Business reasons for starting a change   &  Described in  \\
        \midrule
        Time to market must be reduced    &  P17, P18, P32, P38, P39, P47  \\
        Delivery is considered slow       &  P2, P8, P25, P36, P46, P49  \\
        Something must be done to remain competitive  &
                P9, P10, P11, P18, P19, P32, P38, P42, P49  \\
        Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
        Need to accommodate change during projects  &
                P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
        \bottomrule
    \end{tabular}
    \caption{Business related reasons for starting an agile transformation}
    \label{table:reasonstochange_business}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        \toprule
        Management problems           &  Described in  \\
        \midrule
        Software development is managed badly   &  P8, P21, P30, P38, P44, P49  \\
        Command and control style management    &  P9, P25, P30, P40, P44  \\
        Scheduling and estimation is difficult  &  P16, P19, P21, P30, P33, P40, P41, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Management problems stated as reasons for organizational change}
    \label{table:reasonstochange_management}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        \toprule
        Problems in the old process           &  Described in  \\
        \midrule
        Excess and overhead in process     &  P9, P24, P35, P36, P38, P40, P46, P49, P51  \\
        The process is being ignored       &  P6, P9, P46, P49  \\
        Late integration, testing, and feedback  &  P5, P25, P33, P40, P50  \\
        New practices can not be integrated in the old process  &  P10, P26  \\
        The old process does not scale up  &  P16, P30, P50, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Problems because of which the old process should be changed}
    \label{table:reasonstochange_process}
\end{table}

@@

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
        \toprule
        Reasons to change           &  Described in  \\
        \midrule
        Teams can not collaborate properly   &  P21, P38  \\
        Silos and internal boundaries in organization   &  P27, P40, P46  \\
        Lack of customer collaboration       &  P4, P26, P40  \\
        Demand for better quality            &  P2, P4, P25, P31, P30, P38, P41, P42  \\
        \bottomrule
    \end{tabular}
    \caption{Other reasons why organizations wanted to change the way of working}
    \label{table:reasonstochange_other}
\end{table}

@@

The second research question was aimed to discover how the process of
transformation can be described.
We provide an answer to Research Question 2 in several parts. First the initial
states of organizations are described. Secondly, we present a model for
categorizing transformation types. In the rest of this section we discuss
typical arrangements that relate to the change process, including how training
and piloting is arranged, how change is lead, and what investments are done.

The fact that every organization is unique creates challenges in giving generic
descriptions on how transformations proceed. Further, publications tended to
provide more focused viewpoints rather than describing the transformation
process itself. Therefore it was not sensible to attempt constructing a detailed
model on how large organizations adopt agile methods.
Table \ref{table:transformation_descriptiontypes} presents types of insights
each primary source provided in relation to Research Question 2.

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
        \toprule
        How transformation is described   &  Primary sources  \\
        \midrule
        Narrative description of transformation process &
                P3, P5, P6, P9, P11, P12, P16, P18,
                P24, P27, P30, P36, P40, P41, P46   \\
        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
                P10, P13, P15, P19, P23, P32, P34,
                P39, P47, P48, P50, P52    \\
        Viewpoints of challenges and success factors  &
                P17, P20, P21, P29, P37, P38, P49   \\
        Discussion on establishing large scale agile practices  &
                P7, P8, P31, P42, P43, P44, P45   \\
        Only brief discussion on transformation process  &
                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
        \bottomrule
    \end{tabular}
    \caption{The theme of each primary source, in relation to providing
             insights of the transformation process}
    \label{table:transformation_descriptiontypes}
\end{table}

@@

Based on the primary studies we identified two dimensions in the characteristics
of transformation processes. The first dimension is the pivot point of
leadership in change. This dimension has two extremes: top-down leadership where
the change is mandated by management, and bottom-up leadership where the chage
emerges from team level. The second dimension is the velocity of the change.
Organizations were reported to perform transformations as a big bang or
overnight fashion, or as a step wise process lasting for a long time. Figure
\ref{fig:transformation_dimensions} summarizes the transformation process
dimensions.

\begin{figure}[h]
  \begin{center}
    \includegraphics{graphics/transformation_dimensions.pdf}
    \caption{Dimensions for categorizing transformation types}
    \label{fig:transformation_dimensions}
  \end{center}
\end{figure}

We do not quantitively measure cases on the transformation process dimensions,
because the criteria for creating a measurement would be vague. For instance, we
do not think that presenting a single quantitative measurement of the weight
point of leadership on an axis between management centric and team centric would
describe reality accurately. The dimensions of pivot of leadership and
transformation velocity nevertheless exist.

As the primary sources describe, transformation cases combined the dimensions of
leadership pivot and velocity. Four combinations of the characteristics could be
observed. Figure \ref{fig:transformation_types} summarizes the combinations of
transformation characteristics.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/transformation_process.pdf}
    \caption{Categorization of transformation types}
    \label{fig:transformation_types}
  \end{center}
\end{figure}

The most extreme case of adoption was a bing bang transformation. To realize
such a large change it is necessary that the change was centrally coordinated.
For step wise transformations there were three variants depending on the amount
of involvement of management and development teams. In some cases teams palyed a
passive role, and in some cases management was described to be reluctant to
change. In many cases both management and devlopment dealed with the
transformation on their respective level of influence.

As a concept separate from the dimensions of leadership pivot and velocity, the
transformation was described as a continuous process. There were a few
viewpoints illustrating the concept of continuous transformation. Several
organizations strived to introduce a culture of continuous improvement, which
would make organizational adjustment a permanent state. In many cases the agile
transformation was not considered to have ended at any distinct point in time,
and the impression that it was still ongoing was given. Finally, transformation
was simply described as hard work that will inevitably take a long time.

% Mentioning big bang versus step wise [P16].

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        \toprule
        Change leaders         &  Described in  \\
        \midrule
        High-up change leaders or sponsors  &
                P3, P5, P24, P31, P36, P38, P46, P49  \\
        Agile task group for coordinating change   &
                P4, P16, P18, P21, P23, P25, P30, P40, P41,
                P43, P45, P48, P49, P50  \\
        Change leaders on other organizational levels  &
                P2, P6, P9, P14, P17, P18, P27  \\
        Informal spreading of agile awareness  &
                P3, P6, P14, P40  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of entities leading change in the agile transformation}
    \label{table:transformation_leadership}
\end{table}

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
        \toprule
        Activity           &  Described in  \\
        \midrule
        Training    &
                P3, P6, P8, P9, P11, P16, P18-P21, P23-P25, P27, P30,
                P32, P35, P37, P39-P41, P43-P46, P48-P51  \\  % 29
        Coaching    &
                P1-P3, P6, P8, P9, P11, P16, P18, P23, P27, P29,
                P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
        In-house coaching   &
                P1-P3, P6, P11, P16, P23, P29, P31, P32,
                P39, P41, P45, P47, P49, P50, P52  \\    % 17
        Seminars and workshops   &
                P1-P4, P7, P9, P10, P18, P11, P16, P23, P25,
                P26, P27, P32, P41, P44, P46, P47   \\   % 19
        Communities, special interest groups  &
                P3, P6, P7, P9, P10, P11, P12, P14, P32, P41,
                P42, P45, P47, P52  \\   % 14 
        \bottomrule
    \end{tabular}
    \caption{Summary of activities for building awareness and knowledgeability
             in the agile way of working}
    \label{table:transformation_training}
\end{table}

@@

A logical first step in an agile transformation is to try out the new methods in
a pilot project. A large amount of the primary studies described how piloting
was used in the beginning of the transformation. Primary sources discussing
the use of piloting are listed in table \ref{table:transformation_piloting}.

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
        \toprule
        Sources discussing viewpoints on piloting   & 
                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
                P36, P38, P40, P41, P43, P45, P46, P47  \\
        \bottomrule
    \end{tabular}
    \caption{Summary of activities aiming to advance the transformation}
    \label{table:transformation_piloting}
\end{table}

@@

Even if any activities aiming for transformation can be considered as
investments in themselves, the transformation process included some specific
tangible investments. These investments were focused on better enabling agile
software development. They also highlight the concerns that organizations had when
implementing an agile model. Typical investments were bringing in external help,
new management tools, improvements in testing environments, and updating of the
physical working spaces. Table \ref{table:transformation_investments} summarizes
common types of investments.

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.35\textwidth}
                     >{\raggedright\arraybackslash}p{0.57\textwidth} }
        \toprule
        Investments during transformation      &  Described in  \\
        \midrule
        Using external consultants  &
                P5-P9, P16-P21, P23 P25, P27, P29-P31,
                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
        New project management tools for agile  &
                P10, P11, P14, P16, P17, P26, P33,
                P41, P43, P44, P45, P48, P50 \\
        Continuous builds and testing  &
                P5, P11, P16, P17, P25, P27, P34, P39, P42, P50 \\
        Proprietary agile material or documentation   & 
                P1, P2, P8, P14, P16, P23, P26, P29, P32, P45  \\
        Refurbishing work areas and other spaces   &
                P6, P22, P35, P36, P38, P42, P46, P48 \\
        \bottomrule
    \end{tabular}
    \caption{Summary of investments done during transformation, aiming to make
             agile methods perform better}
    \label{table:transformation_investments}
\end{table}
